---
title: "HackCancer"
author: "RLadies"
date: "10 June 2018"
output: 
  html_document:
    toc: true
    toc_depth: 4
    code_folding: hide
    df_print: paged
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(dplyr)
```

## COSMIC Dataset

COSMIC dataset contains information about a list of 1.397.048 samples.

```{r load COSMIC}
samples <- read.csv('~/HackCancer/data/COSMIC/Cosmicsample.tsv',
              sep = "\t", na.strings = c("NS", ""))

samples_summary <- summary(samples)

samples <- samples %>%
  dplyr::mutate(Histology = ifelse(Histology.subtype.1 == "NS", 
                                   as.character(Primary.histology), 
                                   as.character(Histology.subtype.1)))
```

### Gender

```{r gender}

gender_summary <- samples %>%
  dplyr::group_by(gender) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::mutate(perc = count / sum(count) * 100)
```

There are quite a lot of missing values regarding gender in samples.

* Female (f): `r gender_summary$count[gender_summary$gender == "f"]` samples, `r round(gender_summary$perc[gender_summary$gender == "f"])` % of total.
* Male (m): `r gender_summary$count[gender_summary$gender == "m"]` samples, `r round(gender_summary$perc[gender_summary$gender == "m"])` % of total.
* Undefined (u): `r gender_summary$count[gender_summary$gender == "u"]` samples, `r round(gender_summary$perc[gender_summary$gender == "u"])` % of total.

Let's see how this proportions change by site or by histology:

```{r gender plots}

samples %>% 
  dplyr::select(gender, age, Primary.site) %>%
  dplyr::group_by(gender, Primary.site) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::group_by(Primary.site) %>% 
  dplyr::mutate(freq = count/sum(count)) %>% 
  dplyr::select(-count) %>% 
  dplyr::ungroup() %>% 
  tidyr::spread(gender, freq) %>%
  dplyr::mutate(f = ifelse(is.na(f), 0, f),
                m = ifelse(is.na(m), 0, m),
                u = ifelse(is.na(u), 0, u)) %>% 
  ggtern::ggtern(ggtern::aes(x = f, y = m, z = u)) +
  ggplot2::geom_point() +
  ggplot2::ggtitle("Gender by sites")

samples %>% 
  dplyr::select(gender, Histology) %>%
  dplyr::filter(!is.na(Histology)) %>% 
  dplyr::group_by(gender, Histology) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::group_by(Histology) %>% 
  dplyr::mutate(freq = count/sum(count)) %>%
  dplyr::select(-count) %>%
  dplyr::ungroup() %>% 
  tidyr::spread(gender, freq) %>%
  dplyr::mutate(f = ifelse(is.na(f), 0, f),
                m = ifelse(is.na(m), 0, m),
                u = ifelse(is.na(u), 0, u)) %>% 
  ggtern::ggtern(ggtern::aes(x = f, y = m, z = u)) +
  ggplot2::geom_point() +
  ggplot2::ggtitle("Gender by histology")
```

We can see that some sites and some histologies are suffered just by one of genders (female or male), while others are suffered by both. Unfortunatly, most of them are undefined.

### Age

Most common ages are not the same for every pathology. Some occur more frequently in chilhood, even if most do in adults. Bimodality is also perceived in some of the sites, regarding frequency along samples's ages.

```{r age}
samples <- samples %>%
  dplyr::filter(age != 136)
```

```{r densities simple}
s <- samples %>% 
  dplyr::filter(!is.na(age)) %>%
  dplyr::group_by(Primary.site) %>% 
  dplyr::summarise(sum_count = n())

pat_samples <- samples %>%
  dplyr::select(Primary.site, age) %>% 
  dplyr::left_join(s, by = "Primary.site") %>% 
  dplyr::filter(sum_count > 10)

pat_samples %>%
  ggplot2::ggplot(ggplot2::aes(x = age, color = Primary.site)) +
  ggplot2::geom_density() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::coord_cartesian(xlim = c(0, 100)) +
  ggplot2::ggtitle("Age by site") +
  ggplot2::ggsave(file = "Age.png", width = 5, heigh = 5)
```

Let's see this from Joy Division's perspective:

```{r densities function}
get.density.df <- function(pat_samples){
  d <- density(pat_samples$age)
  data.frame(age = d$x, density = d$y)
}
```

```{r densities joy division function}
joy.division.plot <- function(k, densities){
  p <- ggplot2::ggplot()
  pathology_loop <- unique(densities$Primary.site)
  i <- length(pathology_loop)*k + k
  for (site in pathology_loop) {
    site_density <- densities %>%
      dplyr::filter(Primary.site == site) %>% 
      dplyr::mutate(Primary.site = as.character(Primary.site),
                    density = i + density)
    p <- p + 
      ggplot2::geom_ribbon(data = site_density, alpha = 0.2,
                           ggplot2::aes(x = age, ymax = density), ymin = i,
                           colour = "white", fill = "black", size = 0.5)
    i <- i - k
  }
  
  p + 
    ggplot2::theme(panel.grid = ggplot2::element_blank(),
                   panel.background = ggplot2::element_rect(fill = "#000000"),
                   axis.text = ggplot2::element_blank(),
                   axis.ticks = ggplot2::element_blank(),
                   axis.title = ggplot2::element_blank()) +
    ggplot2::geom_rect(ggplot2::aes(ymin = 0, ymax = k, xmin = -Inf, xmax = Inf),
                       color = "black", fill = "black") 
  
}
```

```{r densities joy division, fig.height=8, fig.align="center"}
pathologies <- pat_samples %>%
  dplyr::select(Primary.site) %>% 
  na.omit() %>% 
  unique() 


densities <- lapply(pathologies$Primary.site,
                    function(p, x){dplyr::filter(x, Primary.site == p) %>% 
                        get.density.df() %>% 
                        dplyr::mutate(Primary.site = p)},
                    dplyr::select(samples, Primary.site, age)) %>% 
  dplyr::bind_rows()


d_order <- densities %>%
  dplyr::group_by(Primary.site) %>%
  dplyr::summarise(q2_age = median(age))

densities <- densities %>%
  dplyr::left_join(d_order, by = "Primary.site") %>% 
  dplyr::arrange(q2_age)

joy.division.plot(0.01, densities) +
  ggplot2::ggtitle("Age by site")
```
